---
title: "Homework 4"
author: "Bahadır Yüzlü"
date: "r"
format: 
  html:
      toc: true
      embed-resources: true
editor: visual
---

IE451 Homework 4

```{r, include=FALSE}
library(tidyverse)
library(magrittr)
library(ISLR2)
library(ggplot2)
library(pander)
library(car)
```

## Question 6

### A
#### Estimate the probablity that a student who studies for 40h and has an undergrad GPA of 3.5 gets an A in the class.

We suppose that we have a model. According to the logistic regression, 
$$
\text{p}(X_1,X_2) = \frac{e^{\beta_0 + \beta_1X_1 + \beta_2X_2 }}{1+e^{\beta_0 + \beta_1X_1 + \beta_2X_2}}
$$
Then by equalizing 
$$\beta_0 = -6, \beta_1 =0.05, \beta_2 = 1  $$
We got
$$\text{p}(X_1,X_2) = \frac{e^{-6 + 0.05X_1 + X_2}}{1+e^{-6 + 0.05X_1 + X_2}} $$
$$\text{Then finding for a sample } X_1 = 40 \text{ and } X_2 = 3.5 $$
$$\text{p}(40,3.5) = \frac{e^{-6 + 0.05*40 + 3.5}}{1+e^{-6 + 0.05*40 + 3.5}} $$
Then we can say that 
$$\frac{e^{-0.5}}{1+e^{-0.5}}= \frac{0.606}{1.606} = 0.377 $$

### B
#### How many hours would the student in part(a) need to study to have a 50% chance of A.

The model is the same, so we can use the odd version of the model for computation easiness.


$$
\log(\frac{\text{p}(X_1,3.5)}{1-\text{p}(X_1,3.5)}) = -6 + 0.05X_1 + 3.5
$$
$$
\frac{\text{p}(X_1,3.5)}{1-\text{p}(X_1,3.5)} = 1
$$
Then we need to adjust the study hour to equalize the system. 
$$
\log{1} = -6 + 0.05X_1 + 3.5
$$
Meaning that;
$$
X_1 = 50
$$
That student should work 50 hours to increase his/her chance up to 50%.

## Question 9

### A

#### On Average, what fraction of people with an odds of 0.37 of defaulting on their credit card payment will in fact default?

First, we need to construct our logistic model to represent the model. 

```{r}
d <- ISLR2::Default
glmod <- glm(default ~ student + balance, data = d, family = binomial)
summary(glmod)
coefficients(glmod)
```
We can construct the model;
$$
\log(odd) = \beta_0 + \beta_1X_1 + \beta_2X_2
$$
$$
\beta_0 = -10.7494, \beta_1 = -0.715, \beta_2 = 0.0057 
$$
If odd is 37%,
$$
\frac{\text{p}(X_1,X_2)}{1-\text{p}(X_1,X_2)} = 0.37
$$
p is appear to be;
$$
\text{p}(X_1,X_2) = \frac{37}{137} = 0.27
$$
Then we should look which of the observations have a prediction probability of 0.27 (I used a %5 interval).
```{r}
possible_default <- d %>%
  mutate(phat = fitted(glmod)) %>% 
  filter(0.2565<phat) %>% 
  filter(phat<0.2835)
possible_default
```
And finally how many of them are defaulted in fact.

```{r}
sum(possible_default$default == "Yes")
```
The answer is 7 out of 35.

### B 

#### Suppose that an individual has a 16% chance of defaulting on her credit card payment. What are the odds that she will default. 

$$
\text{p}(X_1,X_2) = \frac{16}{116} = 0.1379
$$

## Question 12

### A

#### What is the log odds of orange versus apple in your model?
$$
\log(odd) = \beta_0 + \beta_1X_1 
$$

### B

#### What is the log odds of orange versus apple in your friend's model?

$$
\log(odd) = (\alpha_{orange,0}-\alpha_{apple,0}) +(\alpha_{orange,1}-\alpha_{apple,1})X_1
$$

### C

#### What are the coefficient estimates in your friend's model?

For my model,
$$
\log(odd) = 2 - X_1 
$$

To equalize them,

$$
2 - X  = (\alpha_{orange,0}-\alpha_{apple,0}) +(\alpha_{orange,1}-\alpha_{apple,1})X_1
$$
Then it could be said that coefficient estimates are following;

$$
\alpha_{orange,0}-\alpha_{apple,0} = 2
$$
$$
\alpha_{orange,1}-\alpha_{apple,1} = -1
$$

### D

#### What are the coefficients in your model?

$$
\log(odd) = (1.2-3) +(-2-0.6)X_1 = (-2.06)X_1 - 1.8
$$
Then to fit my model into that, I arrange my estimators like following;

$$
\beta_0 = -1.8, \beta_1 =-2.06  
$$

### E

#### Finally, what friction of time do you expect agreement between the 2 model?  

Actually, just equalizing odd logarithm to determine estimators does not make the two model so alike. Therefore, I expect a low agreement situation between the two.






